---
format:
  revealjs:
    theme: [default, custom.scss]
    logo: "../images/metrics_hex.png"
    footer: "[ECON 480 — Econometrics](https://metricsF22.classes.ryansafner.com)"
    height: 900
    width: 1600
    df-print: paged
    slide-number: c
    chalkboard: true
overview: true
execute:
  echo: false
  warning: false
  freeze: auto
---

##  {data-menu-title="Title Slide" background-image="images/metrics_title_slide.png"}

[Midterm Review]{.custom-title}

[ECON 480 • Econometrics • Fall 2022]{.custom-subtitle}

[Dr. Ryan Safner <br> Associate Professor of Economics]{.custom-author}

[<a href="mailto:safner@hood.edu"><i class="fa fa-paper-plane fa-fw"></i>safner\@hood.edu</a> <br> <a href="https://github.com/ryansafner/metricsF22"><i class="fa fa-github fa-fw"></i>ryansafner/metricsF22</a><br> <a href="https://metricsF22.classes.ryansafner.com"> <i class="fa fa-globe fa-fw"></i>metricsF22.classes.ryansafner.com</a><br>]{.custom-institution}

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(broom)
library(haven)
library(kableExtra)
library(patchwork)
library(fontawesome)
library(gapminder)
library(ggthemes)
library(scales)
library(infer)
library(modelsummary)
knitr::opts_chunk$set(echo=F,
                      message=F,
                      warning=F)
update_geom_defaults("label", list(family = "Fira Sans Condensed"))
update_geom_defaults("text", list(family = "Fira Sans Condensed"))

```

## Question 1

What does *endogenous* mean, in words? What about statistically?

## Question 2

If a regression is *biased* (from endogeneity), what can we learn about the bias?

## Question 3

What does *heteroskedasticity* mean? Does *heteroskedasticity* **bias** $\hat{\beta_1}$?

## Question 4

Is this data likely *heteroskedastic* or *homoskedastic*?

```{r, fig.align="center", fig.width=12}
library(tidyverse)
het_data<-tibble(x = runif(500,-3,3),
                 y = rnorm(500, 0, sd = 2 + x^2))

ggplot(data = het_data)+
  aes(x = x,
      y = y)+
  geom_point(color="blue")+
  geom_smooth(method="lm", color="red")+
    labs(x = "X",
       y = "Y")+
  ggthemes::theme_pander(base_family = "Fira Sans Condensed",
           base_size=20)
```

## Question 5

What three things impact the variation of $\hat{\beta_1}$? How?


## Question 6

What are the four assumptions we make about the error term?

. . .

Which is most important?

## Question 7

$$Wages_i=\beta_0+\beta_1Education+u_i$$

. . .

a. What is in $u_i$?

. . .

b. Is $\hat{\beta_1}$ likely biased?

## Question 8

What does $R^2$ measure? What does it mean? How do we calculate it?

## Question 9

What does $\sigma_u$ (SER) measure? What does it mean?

## Question 10

Interpret all of these numbers (except Adjusted R-squared and the last line): 
```{r}
summary(lm(y~x, data = het_data))
```

## Question 11

Interpret all of these numbers: 

```{r}
reg<-lm(y~x, data = het_data)
modelsummary(models = list("y" = reg),
             fmt = 2, # round to 2 decimals
             output = "html",
             coef_rename = c("(Intercept)" = "Constant",
                             "str" = "STR"),
             gof_map = list(
               list("raw" = "nobs", "clean" = "n", "fmt" = 0),
               list("raw" = "r.squared", "clean" = "R<sup>2</sup>", "fmt" = 2),
               #list("raw" = "adj.r.squared", "clean" = "Adj. R<sup>2</sup>", "fmt" = 2),
               list("raw" = "rmse", "clean" = "SER", "fmt" = 2)
             ),
             escape = FALSE,
             stars = c('*' = .1, '**' = .05, '***' = 0.01)
)
```

## Question 12

Suppose $Y$ is normally distributed with a mean of 10 and a standard error of 5. What is the probability that $Y$ is between 5 and 15?


## Question 13

Explain what a $Z$-score means.

## Question 14

Explain what a $p$-value means. 

## Question 15

We run the following hypothesis test at $\alpha=0.05$: 

$$\begin{align*}
H_0: \, & \beta_1=0\\
H_1: \, & \beta_1 \neq 0 \\
\end{align*}$$

Is this test one-sided or two-sided? 

We find the $p$-value is 0.02. What is our conclusion? Be specific and precise in your wording!

## Question 16

Suppose we ran that hypothesis test on our finding. What can we conclude?

::: columns
::: {.column width="50%"}
```{r}
summary(lm(y~x, data = het_data))
```
:::
::: {.column width="50%"}
```{r}
modelsummary(models = list("y" = reg),
             fmt = 2, # round to 2 decimals
             output = "html",
             coef_rename = c("(Intercept)" = "Constant",
                             "str" = "STR"),
             gof_map = list(
               list("raw" = "nobs", "clean" = "n", "fmt" = 0),
               list("raw" = "r.squared", "clean" = "R<sup>2</sup>", "fmt" = 2),
               #list("raw" = "adj.r.squared", "clean" = "Adj. R<sup>2</sup>", "fmt" = 2),
               list("raw" = "rmse", "clean" = "SER", "fmt" = 2)
             ),
             escape = FALSE,
             stars = c('*' = .1, '**' = .05, '***' = 0.01)
)
```
:::
:::
