---
title: "1.2 — Meet R — Class Content"
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance=minimal}
# Meeting Dates
*Thursday, August 26, 2021*
:::

::: .{callout-warning}
# Upcoming Assignment

Please complete the [preliminary survey](/assignment/00-preliminary-survey).
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we begin the long slog to your mastery of `R`. We begin with the basics - how R works, how to use it, the different data types, and how to create and manipulate objects.

## <i class="fas fa-book-reader fa-lg"></i> Readings

- <i class="fas fa-external-link-square-alt"></i>Ch.1 in Wickham & Grolemund, [http://r4ds.had.co.nz/introduction.html](R for Data Science)

Now that we start working with R, you should consider this book to be your *primary* reference for R-related questions. We will broadly cover the first few chapters in order over the next 2-3 class periods.

## <i class="fas fa-registered"></i> R Practice

Today we will be working on [practice problems](/r/1.2-r). Answers will be posted later on that page.

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides. 

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("1.2-slides")
```

## <i class="fas fa-laptop-code"></i> Assignments

### Preliminary Statistics Survey

Please take the [preliminary survey](/assignment/00-preliminary-survey) on your statistics and software background. This will help us all have a productive semester together.
