---
title: "1.4 — Data Wrangling — Class Content"
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance="minimal"}
# Meeting Dates

*Wednesday, September 7, 2022*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we will cover the heart of `tidyverse` and use it for ["data wrangling"](https://en.wikipedia.org/wiki/Data_wrangling). Today will again be *a lot* of content thrown at you, so you can look back at this as a reference all semester. Then we will do more practice problems.

## <i class="fas fa-book-reader fa-lg"></i> Readings

- <i class="fas fa-external-link-square-alt"></i> Chs. 5, 10-12, 18 in Wickham & Grolemund, [R for Data Science](http://r4ds.had.co.nz/)

### References by Package/Task

- `tibble`
  - [*R For Data Science*, Chapter 10: Tibbles](https://r4ds.had.co.nz/tibbles.html)
- `readr` and importing data
  - [*R For Data Science*, Chapter 11: Data Import](https://r4ds.had.co.nz/data-import.html)
  - [R Studio Cheatsheet: Data Import](https://www.rstudio.com/resources/cheatsheets/#import)
- `dplyr` and data wrangling
  - [*R For Data Science*, Chapter 5: Data Transformation](https://r4ds.had.co.nz/tibbles.html)
  - [R Studio Cheatsheet: Data Wrangling](https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf) ([New version](https://www.rstudio.com/resources/cheatsheets/#dplyr))
- `tidyr` and tidying or reshaping data
  - [*R For Data Science*, Chapter 12: Tidy Data](https://r4ds.had.co.nz/tidy-data.html)
  - [R Studio Cheatsheet: Data Wrangling](https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf)
  - [R Studio Cheatsheet: Data Import](https://www.rstudio.com/resources/cheatsheets/#import)
- joining data
  - [*R For Data Science*, Chapter 13: Relational Data](https://r4ds.had.co.nz/relational-data.html)
  - [R Studio Cheatsheet: Data Transformation](https://www.rstudio.com/resources/cheatsheets/#dplyr)

## <i class="fas fa-registered"></i> R Practice

Answers from last class' practice problems on base R are posted [on that page](/r/1.3-r).

Today you will be working on [R practice problems](/r/1.4-r) on data wrangling Answers will be posted later on that page.

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides.

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("1.4-slides")
```
