---
title: "1.5 — Optimize Workflow — Class Content"
execute:
  freeze: auto

---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance="minimal"}
# Meeting Dates

*Monday, September 12, 2022*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today I will give you about half the class period to finish your [1.4 practice problems](/r/1.5-r) and I can answer questions. Then I will briefly show you how you the magic of how you can improve your workflow, efficiency, automation, reproducibility, and safe file backups with R Markdown, R Projects, git and github. All of these tools are slightly advanced, and *optional* for you to use, but the learning curve pays off very high returns for the rest of your life! A good handful of people every semester tell me that they keep using these tools years in the future.

## <i class="fas fa-book-reader fa-lg"></i> Readings

- <i class="fas fa-external-link-square-alt"></i> Chs. 8, 27, 28, 29, 30 (but especially 27) in Wickham & Grolemund, [R for Data Science](http://r4ds.had.co.nz/)

### On Plain-Text Workflow

- Healey, Kieran, [*The Plain Person’s Guide to Plain Text Social Science*](http://plain-text.co/)

### On Markdown Syntax

- [Markdown tutorial](https://www.markdowntutorial.com/)
- Quarto Documentation: [Markdown Basics](https://quarto.org/docs/authoring/markdown-basics.html)

### On Quarto
- Quarto Documentation: [Tutorial: Hello, Quarto](https://quarto.org/docs/get-started/hello/rstudio.html)
- Quarto Documentation: [Tutorial: Computation](https://quarto.org/docs/get-started/computations/rstudio.html)
- Quarto Documentation: [Tutorial: Authoring](https://quarto.org/docs/get-started/authoring/rstudio.html)
- Quarto Documentation: [Guide](https://quarto.org/docs/guide/)

### On LaTeX

- [Wikibooks LaTeX](https://en.wikibooks.org/wiki/LaTeX/) (see esp. [Mathematics](https://en.wikibooks.org/wiki/LaTeX/) chapter)

### On Git and Github

- Bryan, Jenny and Jim Hester, [*Happy Git and GitHub for the useR*](https://happygitwithr.com/)

## <i class="fas fa-registered"></i> R Practice

Answers from last class' practice problems on base R are posted [on that page](/r/1.3-r).

Today you will be working on [R practice problems](/r/1.4-r) on data wrangling Answers will be posted later on that page. 

Today’s "[practice problems](/r/1.5-r)" get you to practice the tools we are working with today. They are again not required, but will help you if you are interested.


## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides.

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("1.5-slides")
```
