---
title: "2.1 — Data 101 & Descriptive Statistics — Class Content"
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance="minimal"}
# Meeting Dates

*Wednesday, September 14, 2022*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we begin with a review and overview of using data and descriptive statistics. We want to quantify characteristics about samples as *statistics*, which we will later use to *infer* things about populations (between which we will later *identify* causal relationships).

Next class will be on random variables and distributions. This full week is your crash course/review of basic statistics that we will need to start the "meat and potatoes" of this class: linear regression next Thursday As such, I'll give you a brief homework next week to review these statistical concepts (with minimal use of `R`!).

## <i class="fas fa-book-reader fa-lg"></i> Readings

- <i class="fas fa-book"></i> Math and Probability Background Appendix A in Bailey

Now that we return to the statistics, we will do a minimal overview of basic statistics and distributions. Review Bailey's appendices. Today, only A is really useful, the rest will come next class.

Chapter 2 is optional, but will give you a good overview of using data.

## <i class="fas fa-sigma"></i> Appendix

See the online appendix for today's content:

- [The Summation Operator](/resources/appendices/2.1-appendix.html)

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides.

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r, eval = T}
#| echo: false
#| output: asis

slide_links("2.1-slides")
```
