---
title: "2.2 — Random Variables & Distributions — Class Content"
execute:
  freeze: auto

---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance="minimal"}
## Meeting Dates

*Wednesday, September 14, 2022*
:::

::: {.callout-warning}
## Upcoming Assignment

[Problem Set 1](/assignment/01-problem-set) is due on Wednesday September 21.

[Problem Set 2](/assignment/02-problem-set) is due on Wednesday September 28.
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we finish your crash course/review of basic statistics with random variables and distributions.

## <i class="fas fa-book-reader fa-lg"></i> Readings

- <i class="fas fa-book"></i> Math and Probability Background Appendix A in Bailey

Now that we return to the statistics, we will do a minimal overview of basic statistics and distributions. Review all of Bailey's appendices.

Chapter 2 is optional, but will give you a good overview of using data.

## <i class="fas fa-sigma"></i> Appendix

See the online appendix for today's content:

- [Properties of Expected Value and Variance](/resources/appendices/2.2-appendix.html#properties-of-expected-value-and-variance)
- [Creating Mathematical Functions in R](/resources/appendices/2.2-appendix.html#creating-mathematical-functions-in-r)
- [Graphing Mathematical and Statistical Functions in R](/resources/appendices/2.2-appendix.html#graphing-mathematical-and-statistical-functions-in-r)

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides.

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r, eval = T}
#| echo: false
#| output: asis

slide_links("2.2-slides")
```
