---
title: "2.3 — Simple Linear Regression — Class Content"
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance="minimal"}
## Meeting Dates

*Wednesday, September 21, 2022*
:::

::: {.callout-warning}
## Upcoming Assignment

[Problem Set 1](/assignment/01-problem-set) is due on Wednesday September 21.

[Problem Set 2](/assignment/02-problem-set) is due on Wednesday September 28.
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we start looking at *associations* between variables, which we will first attempt to quantify with measures like *covariance* and *correlation*. Then we turn to fitting a line to data via *linear regression*. We overview the basic regression model, the parameters and how they are derived, and see how to work with regressions in `R` with `lm` and the tidyverse package [`broom`](https://broom.tidyverse.org/).

We consider an extended example about class sizes and test scores, which comes from a (Stata) dataset from an old textbook that I used to use, [Stock and Watson, 2007](https://www.pearson.com/us/higher-education/product/Stock-Introduction-to-Econometrics-Update-3rd-Edition/9780133486872.html?tab=resources). Download and follow along with the data from today's example:^[Note this is a `.dta` Stata file. You will need to (install and) load the package `haven` to `read_dta()` Stata files into a dataframe.]

- [<i class="fas fa-table"></i> `caschool.dta`](http://metricsf22.classes.ryansafner.com/files/data/caschool.dta)

I have also made a RStudio Cloud project documenting all of the things we have been doing with this data that may help you when you start working with regressions (next class):

- [<i class="fas fa-cloud"></i> Class Size Regression Analysis](https://rstudio.cloud/spaces/270168/content/4559643)

## <i class="fas fa-book-reader fa-lg"></i> Readings

- <i class="fas fa-book"></i> Ch. 3.1, Math and Probability Background Appendix A in Bailey

Now that we return to the statistics, we will do a minimal overview of basic statistics and distributions. Review all of Bailey's appendices.

Chapter 2 is optional, but will give you a good overview of using data.

## <i class="fas fa-sigma"></i> Appendix

See the online appendix for today's content:

- [Variance](/resources/appendices/2.3-appendix.html#variance)
- [Covariance](/resources/appendices/2.3-appendix.html#covariance)
- [Correlation](/resources/appendices/2.3-appendix.html#correlation)

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides.

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r, eval = T}
#| echo: false
#| output: asis

slide_links("2.3-slides")
```
