---
title: "2.4 — Goodness of Fit and Bias — Class Content"
execute:
  freeze: auto

---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance="minimal"}
## Meeting Dates

*Monday, September 26, 2022*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we continue looking at basic OLS regression. We will cover how to measure if a regression line is a good fit (using $R^2$ and $\sigma_u$ or SER), and whether OLS estimators are *biased*. These will depend on four critical *assumptions about $u$.* 

In doing so, we begin an ongoing exploration into inferential statistics, which will finally become clear in another week. The most confusing part is recognizing that there is a *sampling distribution of each OLS estimator*. We want to measure the center of that sampling distribution, to see if the estimator is *biased*. Next class we will measure the spread of that distribution. 

We continue our extended example about class sizes and test scores, which comes from a (Stata) dataset from an old textbook that I used to use, [Stock and Watson, 2007](https://www.pearson.com/us/higher-education/product/Stock-Introduction-to-Econometrics-Update-3rd-Edition/9780133486872.html?tab=resources). Download and follow along with the data from today's example:[^1]

[^1]: Note this is a `.dta` Stata file. You will need to (install and) load the package `haven` to `read_dta()` Stata files into a dataframe.

-   [<i class="fas fa-table"></i> `caschool.dta`](http://metricsf22.classes.ryansafner.com/files/data/caschool.dta)

I have also made a RStudio Cloud project documenting all of the things we have been doing with this data that may help you when you start working with regressions (next class):

-   [<i class="fas fa-cloud"></i> Class Size Regression Analysis](https://rstudio.cloud/spaces/270168/content/4559643)

## <i class="fas fa-book-reader fa-lg"></i> Readings

-   <i class="fas fa-book"></i> Ch. 3.2-3.4, 3.7-3.8 in Bailey, *Real Econometrics*

## <i class="fas fa-sigma"></i> Appendix

See the online appendix for today's content:

- [Deriving the OLS Estimators](/resources/appendices/2.4-appendix.html#deriving-the-ols-estimators)
- [Bias in $\hat{\beta_1}$](/resources/appendices/2.4-appendix.html#bias-in-hatbeta_1)
- [Proof of the Unbiasedness of
 $\hat{\beta_1}$](/resources/appendices/2.4-appendix.html#proof-of-the-unbiasedness-of-hatbeta_1)

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides.

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r, eval = T}
#| echo: false
#| output: asis

slide_links("2.4-slides")
```
