---
title: "2.5 — Precision and Diagnostics — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance="minimal"}
## Meeting Dates

*Wednesday, September 28, 2022*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Last class and this class we are looking at the *sampling distibution* of OLS estimators (particularly $\hat{\beta_1})$. Last class we looked at what the *center* of the distribution was - the true $\beta_1$ - so long as the assumptions about $u$ hold:

- When $cor(X,u)=0$, $X$ is *exogenous* and the OLS estimators are *unbiased*.
- What $cor(X,u)\neq 0$, $X$ is *endogenous* and the OLS estimators are *biased*.

Today we continue looking at the *sampling distibution* by determining the variation in $\hat{beta_1}$ (it's variance or its standard error^[The square root of variance, as always!]). We look at the formula and see the three major determinants of variation in $\hat{\beta_1}$:

1. Goodness of fit of the regression $(SER$ or $\hat{\sigma_u}$
2. Sample size $n$
3. Variation in $X$

We also look at the diagnostics of a regression by looking at its residuals $(\hat{u_i})$ for anomalies. We focus on the problem of *heteroskedasticity* (where the variation in $\hat{u_i])$ changes over the range of $X$, which violates assumption 2 (errors are homoskedastic): how to detect it, test it, and fix it with some packages. We also look at outliers, which can bias the regression. Finally, we also look at how to present regression results. 

We continue our extended example about class sizes and test scores, which comes from a (Stata) dataset from an old textbook that I used to use, [Stock and Watson, 2007](https://www.pearson.com/us/higher-education/product/Stock-Introduction-to-Econometrics-Update-3rd-Edition/9780133486872.html?tab=resources). Download and follow along with the data from today's example:[^1]

[^1]: Note this is a `.dta` Stata file. You will need to (install and) load the package `haven` to `read_dta()` Stata files into a dataframe.

-   [<i class="fas fa-table"></i> `caschool.dta`](http://metricsf22.classes.ryansafner.com/files/data/caschool.dta)

I have also made a RStudio Cloud project documenting all of the things we have been doing with this data that may help you when you start working with regressions (next class):

-   [<i class="fas fa-cloud"></i> Class Size Regression Analysis](https://rstudio.cloud/spaces/270168/content/4559643)

## <i class="fas fa-book-reader fa-lg"></i> Readings

-   <i class="fas fa-book"></i> Finish Ch.3 in Bailey, *Real Econometrics*

## <i class="fas fa-registered"></i> R Practice

Today (and next class) you will be working on [practice problems](/r/2.5-r). Answers will be posted on that page later.

## <i class="fas fa-sigma"></i> Appendix

See the online appendix for today's content:

-   [Deriving the OLS Estimators](/resources/appendices/2.4-appendix.html#deriving-the-ols-estimators)
-   [Bias in $\hat{\beta_1}$](/resources/appendices/2.4-appendix.html#covariance)
-   [Correlation](/resources/appendices/2.3-appendix.html#correlation)

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides.

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r, eval = T}
#| echo: false
#| output: asis

slide_links("2.5-slides")
```
