---
title: "2.7 — Hypothesis Testing for Regression — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance="minimal"}
## Meeting Dates

*Wednesday, October 5, 2022*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we finish our discussion of statistical inference, focusing on hypothesis testing about our regression. We will hopefully have time at the end to work on a few practice problems.

## <i class="fas fa-book-reader fa-lg"></i> Readings

- <i class="fas fa-book"></i> Ch.4 in Bailey, *Real Econometrics*
- <i class="fas fa-book"></i> Ch.4 in Bailey, *Real Econometrics*
- [<i class="fas fa-external-link-square-alt"></i> Ch. 10 (optionally 8-9) in *Modern Dive* ](https://moderndive.netlify.com/10-inference-for-regression.html)
- [<i class="fas fa-external-link-square-alt"></i> Visualizing Probability and Inference ](https://seeing-theory.brown.edu/frequentist-inference/index.html)
- [<i class="fab fa-youtube"></i> *Last Week Tonight with John Oliver*: Scientific Studies](https://www.youtube.com/watch?v=0Rnq1NpHdmw)
- [“There is Still Only One Test”](https://allendowney.blogspot.com/2016/06/there-is-still-only-one-test.html)

The readings are the same as last class, plus one highly recommended additional video: be sure to watch the excellent and hilarious discussion of the limits and misuses of scientific studies and statistical significance $(p$-values) in the *Last Week Tonight* clip. I also added the blog post that I reference in class about all hypothesis testing is really one common procedure.

## <i class="fas fa-registered"></i> R Practice

Today we be working on [practice problems](/r/2.7-r). Answers will be posted on that page later.

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides.

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r, eval = T}
#| echo: false
#| output: asis

slide_links("2.7-slides")
```
