---
title: "4.3 — Categorical Data & Interactions — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance="minimal"}
## Meeting Dates

*Monday, November 7, 2022*
*Wednesday, November 9, 2022*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

This week, we look at how to use data that is categorical (i.e. variables that indicate an observation's membership in a particular group or category). We introduce them into regression models as **dummy variables** that can equal 0 or 1: where 1 indicates membership in a category, and 0 indicates non-membership.

We also look at what happens when categorical variables have more than two values: for regression, we introduce a dummy variable for each possible category - but be sure to leave out one reference category to avoid the dummy variable trap.

We then continue by examining how to use categorical data in regression, particularly focusing on *interactions* between variables. We look at three types of interaction effects:
1. Interaction between a continuous variable & a dummy variable
2. Interaction between two dummy variables
3. Interaction between two continuous variables

We will also be working on [practice problems](/r/4.3-r) in `R`.

## <i class="fas fa-book-reader fa-lg"></i> Readings

- <i class="fas fa-book"></i> Ch. 6.1—6.2 in Bailey, *Real Econometrics*

## <i class="fas fa-registered"></i> R Practice

This week you will be working on [R practice problems](/r/4.3-r) on categorical data. Answers will be posted later on that page.

## <i class="fas fa-laptop-code"></i> Assignments

### Problem Set 4 Due Fri Nov 11
[Problem Set 4](/assignments/04-problem-set) is due by the end of the day on Tuesday, November 11.

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides.

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r, eval = T}
#| echo: false
#| output: asis

slide_links("4.3-slides")
```
