---
title: "5.1 — Panel Data and Fixed Effects — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance="minimal"}
## Meeting Dates

*Monday, November 14, 2022* *Wednesday, November 16, 2022*
:::

::: callout-warning
## Upcoming Assignment

[Problem Set 5](/assignment/05-problem-set) is due on Monday November 28.
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today, we begin our brief look at panel data, where we track multiple individuals over time. Panel data contains its own unique challenges, because it contains a time series component for every individual, giving potential sources of bias. 

We now need to understand the third assumption about $u_i$: no autocorrelation. The errors of our observations are likely going to be correlated within each individual and within each time period. 

We can correct for these with a **fixed effects** model that isolates and absorbs some of that bias. In general, for a **two-way** fixed effects model: 

$$\widehat{\text{Y}}_{it} = \beta_0+\beta_1 \text{X}_{1it} + \beta_2 \text{X}_{2it}+\alpha_{i} + \theta_{t} + \nu_{it}$$

Each observation is an individual $i$ at time $t$ (pay attention to the subscripts). 

- Let $Y_{it}$ be our dependent variable, and $X_{1it}$ be the independent variable of interest. We would like to estimate the causal effect of $X_{1it} \rightarrow Y_{it}$.
- $\alpha_i$ is the **group fixed-effect**. It absorbs all unobservable factors that **vary by group** but **don’t change over time**.
- $\tau_t$ is the **time fixed-effect**. It absorbs all unobservable factors that **do not vary by group** but **change over time**.
- Since the fixed effects *do not* pick up factors that **_both_ vary by group _and_ change over time**, we need to include other variables that might cause $X_{1it}$ to be endogenous: hence, $X_{2it}$
- $\nu_{it}$ is the remaining (random) error term (after we have pulled $\alpha_i$ and $\tau_t$ out of $u_{it}$. 

## <i class="fas fa-book-reader fa-lg"></i> Readings

- <i class="fas fa-book"></i> Ch. 8.1—8.4 in Bailey, *Real Econometrics*

## <i class="fas fa-laptop-code"></i> Assignments

### Problem Set 5 Due Mon Nov 28

[Problem Set 5](/assignments/05-problem-set) is due by the end of the day on Wednesday, November 28.

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides.

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r, eval = T}
#| echo: false
#| output: asis

slide_links("5.1-slides")
```
