---
title: "5.3 — Instrumental Variables — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance="minimal"}
## Meeting Dates

*Wednesday, November 30, 2022*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today, we look at another of the most commonly used by professional researchers and econometricians: instrumental variables.

## <i class="fas fa-book-reader fa-lg"></i> Readings

- <i class="fas fa-book"></i> Ch. 9 in Bailey, *Real Econometrics*
- [<i class="fas fa-file-pdf"></i> Levitt, 2002, “Using Electoral Cycles in Police Hiring to Estimate the Effect of Police on Crime: Reply”](/files/readings/Levitt-2002.pdf)
<!--## <i class="fas fa-laptop-code"></i> Assignments

### Problem Set 5 Due Mon Nov 28

[Problem Set 5](/assignments/05-problem-set) is due by the end of the day on Wednesday, November 28.-->

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides.

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r, eval = T}
#| echo: false
#| output: asis

slide_links("5.3-slides")
```
